On the power of oritatami cotranscriptional folding with unary bead sequence\thanks{This work is supported in part by KAKENHI Grant-in-Aid for Challenging Research (Exploratory) No.~18K19779 and JST Program to Disseminate Tenure Tracking System No.~6F36 granted to S.~Z.~F. and S.~S.}

### 2018-12-05 tt_theorem.tex
12/5 morning

\begin{proof}
We consider cases which $w[i-1]$ is two distances away from the singular point.
Let us show that $\#bc(C_{i-1}) > \#bc(C_i)$, that is, when $w[i]$ is stabilized, $w[i]$ uses at least two hands.
Fig.\ref{TTT_a3_w} exhibits all the three kinds of possibility of stabilized $w[i]$.
Suppose $w[i]$ were able to be stabilized with using one hand in each cases.

% Then, $w[i]$ can be also stabilized at $n_3$.

\begin{paragraph}{Case of straight}
\begin{itemize}
\item[-] Case of $n_3$ is free\\
The bead $w[i-2]$ must not have bonds because $w[i]$ is stabilized deterministically so that five neighbors of $w[i-2]$ must be occupied.
Therefore $n_5$ must be occupied.
We can continue this argument in clockwise manner to fill the neighbors of $n_3$ until $n_1$.
However, two neighbors of $n_1$ must be free because they are $n_3$ and the point of $w[i]$, contradict.

\item[-] Otherwise\\
Both $n_3$ and $n_4$ are occupied.
One of $n_1$ and $n_2$ must be free because $S[i] = b$.
We assume $n_1$ is free, then all neighbors of $n_1$ must be occupied because $n_3$ is occupied.
However, one of neighbors $n_1$ is a point which $w[i]$ is stabilized, contradict.
\end{itemize}

Therefore, this case contradict assumption.
\end{paragraph}

\begin{paragraph}{Case of obtuse}
\begin{itemize}
\item[-] Case of $n_3$ is free\\
Any neighbors of $n_3$ have to be occupied but the point which $w[i]$ is stabilized is free. Thus $n_3$ has to be occupied.
\item[-] Case of $n_4$ is free\\
The point $n_2$ must be occupied because $n_4$ is free and $n_5$ must be occupied.
Therefore $n_0$ is occupied.
The bead $w[i]$ forms just one bond so that one of $n_0, n_3$ leave some hands or both of them do not leave any bonds.\\
If $n_0$ has some hands, then $n_3$ does not have any hands so that $n_{-3}$ is occupied. 
We can continue this argument in clockwise manner to fill the neighbors of $w[i]$.
That is, $S[i] = \blacksquare$, contradict.

If $n_3$ has some hands, then $n_0$ does not have any hands so that neighbors of $w[i]$ are occupied. In the same previous way, contradict.\\
If both of $n_0, n_3$ do not have any hands, then both of $n_{-1}, n_{-3}$ are occupied. If one of $n_{-1}, n_{-3}$ has some hands, the other does not have any hands so that $n_{-2}$ is occupied. If both of $n_{-1}, n_{-3}$ do not have any hands, $n_{-2}$ has to be occupied and $n_{-2}$ has some hands, contradict.
\item[-] Case of $n_2$ is free\\
Any neighbors of $n_2$ have to be occupied so that $n_0$ is occupied. Any neighbors of $n_0$ except $n_2$ have to be also occupied but the point which is stabilized $w[i]$ is free. Thus $n_2$ has to be occupied.
\item[-] Case of $n_0$ is free\\
Any neighbors of $n_0$ have to be occupied so that $n_{-1}$ is occupied. Any neighbors of $n_{-1}$ except $n_0$ have to be also occupied but the point which is stabilized $w[i]$ is free. Thus $n_0$ has to be occupied. 
\end{itemize}
\end{paragraph}

\begin{paragraph}{Case of acute}
\begin{itemize}
\item[-] Case of $n_4$ is free\\
The point $n_4$ and a point which is stabilized $w[i]$ are free so that  $w[i-2]$ has some bonds.
However, $w[i]$ can be also stabilized $n_4$ in this case. Thus, $n_4$ has to be occupied.
\item[-] Case of $n_2$ is free\\
The point $n_0$ has to be occupied. The point $n_1$ has to be also occupied because five neighbors of $n_0$ must be occupied.
We consider this case just like case of obtuse and that $n_4$ is free. Then if $w[i-2]$ binds $w[i]$, any $n_{-1}, n_{-2}, n_{-3}$ are occupied. If $n_1$ binds $w[i]$, this case is same. Also if $n_1$ and $w[i-2]$ do not have any hand, any $n_{-1}, n_{-2}, n_{-3}$ are occupied. Therefore, $w[i+1]$ cannot be provided.
\item[-] Case of $n_0$ is free\\
Any neighbors of $n_0$ have to be occupied so that $n_{1}$ is occupied. Any neighbors of $n_{1}$ except $n_0$ have to be also occupied but the point which is stabilized $w[i]$ is free. Thus $n_0$ has to be occupied. 
\item[-] Case of $n_1$ is free\\
Any neighbors of $n_1$ have to be occupied so that $n_{-1}$ is occupied. Any neighbors of $n_{-1}$ except $n_1$ have to be also occupied but the point which is stabilized $w[i]$ is free. Thus $n_1$ has to be occupied. 
\end{itemize}

Therefore, any situations contradict $S[i] = b$.
\end{paragraph}\\

Hence, $w[i]$ uses at least two bonds if $w[i-1]$ is at least two distances away the singular point.
The size of $ \hexagon^2_0 = 19$ and this number contain the singular point, its successor and its three neighbors which is it binds to. Therefore 14 times  $w[i]$ can be stabilized by forming one bond.
Therefore, a deterministic unary oritatami system at $\delta = 1$ and $\alpha = 3$ can yield only finite conformation of length is $2n + 14$.
\end{proof}

>>>>

\begin{proof}
In this proof, we shall verify the claim that when the bead $w[i]$ is stabilized with $S[i] = b$ and $S[i+1] \neq \blacksquare$, if the circle of radius 2 centered at its predecessor $w[i-1]$ is free from the singular point, then $w[i]$ must form at least 2 bonds. 
Recall that the circle of radius 2 centered at the origin $O$, where the only candidate of singularity is, consists of 19 points including $O$. 
In order for the bead at $O$ to be singular, its $\alpha+1 = 4$ neighbors must be occupied. 
This means that there are at most 14 points where a bead find a singular bead within 2 points. 
Therefore, the claim, once proved, and the Tunnel Troll Theorem imply that all but at most 14 beads strictly decrease the binding capability. 
The binding capability of the seed is at most $3n$. 
Consequently, this theorem holds.

Now let us verify the claim. 
Suppose $w[i]$ were stabilized by just one bond. 
There are three cases to be considered as depicted in Fig.~\ref{TTT_a3_w}, depending on the relative position of $w[i]$ to $w[i-2]$ and $w[i-1]$. 
Since $S[i] = b$, at least one of the four neighbors of $w[i-1]$ must be empty. 
If $w_3$ is free, then $w[i-2]$ must have used up all of its hands; otherwise, $w[i]$ would be stabilized also at $n_3$ nondeterministically, a contradiction. 
Thus, $\alpha + 2 = 5$ neighbors of $w[i-2]$, that is, all of its neighbors, must be occupied. 
Hence, $n_5$ is occupied. 
(All the remaining arguments are based on this ``merry-go-round'' occupation. This works only if the circle of radius 2 around $w[i-1]$ is free from the singular bead). 
In the same way, all the neighbors of $n_3$ turned out to be occupied in the clockwise order, but eventually, we would encounter a neighbor that is adjacent to also the point where $w[i]$ is supposed to go. 
Thus, $n_3$ must be occupied (in the left and middle cases). 
In the left case, $n_4$ is symmetric to $n_3$, and hence, it must be occupied, too. 
Since $S[i] = b$, $n_1$ or $n_2$ must be free; assume $n_1$ is. 
Going clockwise around $n_1$ implies that $n_{-1}$ is occupied, but the bead at $n_1$ has a free hand and would cause a nondeterminism.

Let us focus on the remaining cases: middle and right. 
Suppose that among the 5 neighbors of $w[i]$ at which $w[i-1]$ is not, only one can be occupied. 
Otherwise, a bead without any hand is found at one of them, and staring from the point, merry-go-round occupies all the neighbors, but then $w[i+1]$ would lose its way to go. 
In the middle case, this means $n_0$ is free. 
Since $n_{-1}$ is free, so must be $n_2$. 
Repeating this, we get that both $n_4$ and $n_5$ are free, but then $w[i-2]$ would have a free hand and a free neighbor, attract $w[i]$, and cause a nondeterminism. 
Even in the right case, the points $n_1, n_0, n_2, n_4$ turn out to be free one after another likewise, but then $w[i-2]$ would have a free hand and neighbor, a contradiction. 
\qed

%We consider cases which $w[i-1]$ is two distances away from the singular point.
%Let us show that $\#bc(C_{i-1}) > \#bc(C_i)$, that is, when $w[i]$ is stabilized, $w[i]$ uses at least two hands.
%Fig.\ref{TTT_a3_w} exhibits all the three kinds of possibility of stabilized $w[i]$.
%Suppose $w[i]$ were able to be stabilized with using one hand in each cases.
%
%% Then, $w[i]$ can be also stabilized at $n_3$. 
%
%\begin{paragraph}{Case of straight}
%\begin{itemize}
%\item[-] Case of $n_3$ is free\\
%The bead $w[i-2]$ must not have bonds because $w[i]$ is stabilized deterministically so that five neighbors of $w[i-2]$ must be occupied.
%Therefore $n_5$ must be occupied.
%We can continue this argument in clockwise manner to fill the neighbors of $n_3$ until $n_1$.
%However, two neighbors of $n_1$ must be free because they are $n_3$ and the point of $w[i]$, contradict.
%
%\item[-] Otherwise\\
%Both $n_3$ and $n_4$ are occupied.
%One of $n_1$ and $n_2$ must be free because $S[i] = b$.
%We assume $n_1$ is free, then all neighbors of $n_1$ must be occupied because $n_3$ is occupied.
%However, one of neighbors $n_1$ is a point which $w[i]$ is stabilized, contradict.
%\end{itemize}
%
%Therefore, this case contradict assumption.
%\end{paragraph}
%
%\begin{paragraph}{Case of obtuse}
%\begin{itemize}
%\item[-] Case of $n_3$ is free\\
%Any neighbors of $n_3$ have to be occupied but the point which $w[i]$ is stabilized is free. Thus $n_3$ has to be occupied.
%\item[-] Case of $n_4$ is free\\
%The point $n_2$ must be occupied because $n_4$ is free and $n_5$ must be occupied.
%Therefore $n_0$ is occupied.
%The bead $w[i]$ forms just one bond so that one of $n_0, n_3$ leave some hands or both of them do not leave any bonds.\\
%If $n_0$ has some hands, then $n_3$ does not have any hands so that $n_{-3}$ is occupied. 
%We can continue this argument in clockwise manner to fill the neighbors of $w[i]$.
%That is, $S[i] = \blacksquare$, contradict.
%
%If $n_3$ has some hands, then $n_0$ does not have any hands so that neighbors of $w[i]$ are occupied. In the same previous way, contradict.\\
%If both of $n_0, n_3$ do not have any hands, then both of $n_{-1}, n_{-3}$ are occupied. If one of $n_{-1}, n_{-3}$ has some hands, the other does not have any hands so that $n_{-2}$ is occupied. If both of $n_{-1}, n_{-3}$ do not have any hands, $n_{-2}$ has to be occupied and $n_{-2}$ has some hands, contradict.
%\item[-] Case of $n_2$ is free\\
%Any neighbors of $n_2$ have to be occupied so that $n_0$ is occupied. Any neighbors of $n_0$ except $n_2$ have to be also occupied but the point which is stabilized $w[i]$ is free. Thus $n_2$ has to be occupied.
%\item[-] Case of $n_0$ is free\\
%Any neighbors of $n_0$ have to be occupied so that $n_{-1}$ is occupied. Any neighbors of $n_{-1}$ except $n_0$ have to be also occupied but the point which is stabilized $w[i]$ is free. Thus $n_0$ has to be occupied. 
%\end{itemize}
%\end{paragraph}
%
%
%
%\begin{paragraph}{Case of acute}
%\begin{itemize}
%\item[-] Case of $n_4$ is free\\
%The point $n_4$ and a point which is stabilized $w[i]$ are free so that  $w[i-2]$ has some bonds.
%However, $w[i]$ can be also stabilized $n_4$ in this case. Thus, $n_4$ has to be occupied.
%\item[-] Case of $n_2$ is free\\
%The point $n_0$ has to be occupied. The point $n_1$ has to be also occupied because five neighbors of $n_0$ must be occupied.
%We consider this case just like case of obtuse and that $n_4$ is free. Then if $w[i-2]$ binds $w[i]$, any $n_{-1}, n_{-2}, n_{-3}$ are occupied. If $n_1$ binds $w[i]$, this case is same. Also if $n_1$ and $w[i-2]$ do not have any hand, any $n_{-1}, n_{-2}, n_{-3}$ are occupied. Therefore, $w[i+1]$ cannot be provided.
%\item[-] Case of $n_0$ is free\\
%Any neighbors of $n_0$ have to be occupied so that $n_{1}$ is occupied. Any neighbors of $n_{1}$ except $n_0$ have to be also occupied but the point which is stabilized $w[i]$ is free. Thus $n_0$ has to be occupied. 
%\item[-] Case of $n_1$ is free\\
%Any neighbors of $n_1$ have to be occupied so that $n_{-1}$ is occupied. Any neighbors of $n_{-1}$ except $n_1$ have to be also occupied but the point which is stabilized $w[i]$ is free. Thus $n_1$ has to be occupied. 
%\end{itemize}
%
%Therefore, any situations contradict $S[i] = b$.
%\end{paragraph}\\
%
%
%Hence, $w[i]$ uses at least two bonds if $w[i-1]$ is at least two distances away the singular point.
%The size of $ \hexagon^2_0 = 19$ and this number contain the singular point, its successor and its three neighbors which is it binds to. Therefore 14 times  $w[i]$ can be stabilized by forming one bond.
%Therefore, a deterministic unary oritatami system at $\delta = 1$ and $\alpha = 3$ can yield only finite conformation of length is $2n + 14$.
